---
title: "HC - Data Prepare"
output:
  pdf_document: default
---


```{r}

require(corrplot)

require(ggplot2)    ## declaratively creating graphics - https://ggplot2.tidyverse.org/ 

require(gridExtra)          ## arrange visualizations using grid 

```


```{r}

#Load the data and initialize necessary variables

medicalData = read.csv("http://www.datadescant.com/stat109/hospvisits.csv")

head(medicalData)

response_df = medicalData['totalexp']  # Y variable
predictors_df = medicalData[, !names(medicalData) %in% "totalexp" ]  # X variables

# Data frame to store the results of the various models
modelResults = setNames(data.frame(matrix(ncol = 7, nrow = 0)), c("Name", "Model", "RMSE", "R2", "MAE", "ACC")) 

summary(medicalData)

# Create train & test data set

set.seed(123)
sample = sample.int(n = nrow(medicalData), 
                     size = floor(.80*nrow(medicalData)), replace = F)
medicalData.Train = medicalData[sample,]
medicalData.Test = medicalData[-sample,]

```

```{r}
# Data exploration

numericVars = which(sapply(medicalData, is.numeric)) #index vector numeric variables
numericVarNames = names(numericVars) #saving names vector for use later on
##cat('There are', length(numericVars), 'numeric variables')

par(mfrow = c(2, 2))

hist(medicalData$totalexp,
main="Total Helath Care Expenses",
xlab="Expenses in Dollars",
xlim=c(0,150000),
col="blue",
freq=TRUE
)

hist(log(medicalData$totalexp),
main="Log of Total Helath Care Expenses",
xlab="Log of Expenses in Dollars",
col="blue",
freq=TRUE
)

qqnorm(medicalData$totalexp, main = "Total Expense Q-Q Plot", xlab = "Theoretical Quantiles", 
       ylab = "Sample Quantiles", plot.it = TRUE, datax = FALSE)
qqline(medicalData$totalexp, datax = FALSE, distribution = qnorm, 
       probs = c(0.25, 0.75), qtype = 7)


qqnorm(log(medicalData$totalexp), main = "Log Total Expense Q-Q Plot", xlab = "Theoretical Quantiles", 
       ylab = "Sample Quantiles", plot.it = TRUE, datax = FALSE)
qqline(log(medicalData$totalexp), datax = FALSE, distribution = qnorm, 
       probs = c(0.25, 0.75), qtype = 7)
       
```

```{r}
# Check for correlation among the variables

cor_numVar = cor(medicalData, use="pairwise.complete.obs") #correlations of all numeric variables
cor_sorted = as.matrix(sort(cor_numVar[,'totalexp'], decreasing = TRUE))

#select only high corelations
CorHigh = names(which(apply(cor_sorted, 1, function(x) abs(x)>0.1)))
cor_numVar = cor_numVar[CorHigh, CorHigh]

corrplot.mixed(cor_numVar, tl.col="black", tl.pos = "lt")

corrplot(cor_numVar, title = "",

            type = "lower", 

            order = "hclust", 

            hclust.method = "centroid",

            tl.cex = 0.8,

            tl.col = "black", 

            tl.srt = 45)

```

```{r}
# More data exploration

# Race Group
p1 = ggplot(data=medicalData[!is.na(medicalData$totalexp),], aes(x=factor(race_grp), y=totalexp))+
        geom_boxplot(col='blue') + labs(x='Race Group',y='Total Expanse') +
        scale_y_continuous(breaks= seq(0, 200000, by=10000))

# Health Status
p2 = ggplot(data=medicalData[!is.na(medicalData$totalexp),], aes(x=factor(srhealth), y=totalexp))+
        geom_boxplot(col='blue') + labs(x='Health Status',y='Total Expanse') +
        scale_y_continuous(breaks= seq(0, 200000, by=10000))

# Marital Status
p3 = ggplot(data=medicalData[!is.na(medicalData$totalexp),], aes(x=factor(marital), y=totalexp))+
        geom_boxplot(col='blue') + labs(x='Marital Status',y='Total Expanse') +
        scale_y_continuous(breaks= seq(0, 200000, by=10000))

# Mental Health
p4 = ggplot(data=medicalData[!is.na(medicalData$totalexp),], aes(x=factor(mntl_hlth), y=totalexp))+
        geom_boxplot(col='blue') + labs(x='Mental Health',y='Total Expanse') +
        scale_y_continuous(breaks= seq(0, 200000, by=10000))

# Doctor Visits
p5 = ggplot(data=medicalData[!is.na(medicalData$totalexp),], aes(x=dr_visits, y=totalexp)) +
  geom_point(col='blue') +
  geom_smooth(method = "lm", se=FALSE, color="black", aes(group=1)) +
  labs(x='Doctor Visits',y='Total Expanse')

# Education
p6 = ggplot(data=medicalData[!is.na(medicalData$totalexp),], aes(x=factor(educ), y=totalexp))+
        geom_boxplot(col='blue') + labs(x='Education',y='Total Expanse') +
        scale_y_continuous(breaks= seq(0, 200000, by=10000))

grid.arrange(p1,p2,p3,p4,nrow = 2)
grid.arrange(p5,p6,nrow = 1)

```


```{r}

```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
